---
title: "index"
author: "Yan Remes"
date: "11/13/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, warning=FALSE, include=FALSE}
library(tidyverse)
library(ggplot2)
knitr::opts_chunk$set(echo = TRUE)
source("./analysis.R")
```

## Introduction + Summary Information

```{r}
df_trends <- read.csv("incarceration_trends.csv")
df_jurisdiction <- read.csv("incarceration_trends_jail_jurisdiction.csv")

recent_avg_black_jail_pop
recent_avg_white_jail_pop
county_with_highest_black_pop_15to64
county_with_highest_white_pop_15to64
change_In_black_pop
```
df_trends holds data from the Death in Custody Reporting Program that is used to 
find the prison population and the prison admission. Each data takes in yearly estimates. 
Also, population and admission splits into gender and race. In addition
the population splits into age categories
df_jurisdiction holds data from the Annual Survey of Jails and Census of Jails.
Each data takes in yearly estimates. Also, population and admission splits into gender and race. 

recent_avg_black_jail_pop shows the average value of black jail population across all the counties for 2018 is 91.81734
recent_avg_white_jail_pop shows the average value of black jail population across all the counties for 2018 is 127.2046
It shows that in 2018 the average number of whites in the jail is higher than the number of blacks in the jail
by 35.38726
county_with_highest_black_pop_15to64 shows the county with the highest population of black people between the age of 
15 and 64 that is New York County
county_with_highest_white_pop_15to64 shows the county with the highest population of white people between the age of 
15 and 64 that is Hoonah-Angoon Census Area
This analysis helps to understand where what are the major counties where 2 races live
change_In_black_pop shows change of black people people between the age of 15 and 64 in New York County starting from 2000 and until 2018.
The result is -78040.
It shows that the population of black people significantly decreased from 2000 to 2018

## Trends Over Time Chart


```{r}
trends_over_time
```
Sorting data by the most populate states in 2018 where black people ages from 15 to 64 live I was 
interested to see the patterns how the population changes over time. Alaska shows stable results in 
the population of black people from 1990 to 2018.
There is significant increase in black population in Arizona from 2005 until 2018. The increase is 
approximately from 125000 to 250000. Looking at Arkansas we can see a stable moderate increase in 
population from 230000 to 270000. Alabama shows a significant increase starting from 1990 and until 2018. 
The population increased from 600000 to 850000. California also shows significant increase from 1350000 to 1700000. 
This trend is important to consider to see the future potential changes in black population

## Variable comparison chart


```{r}
variable_comparison
```
I used this chart to see the difference in population over time for races in the most populate states in 2018.
Alaska shows stable results from 1990 to 2018 for both race. Alabama shows an increase in population
of black people from 600000 to 850000, while there is a slight decrease in population of white people from 220000 to 200000.
Arizona shows slight increase in black population but it is not very sufficient, white population shows a slight decrease
but it is also not very sufficient. Arizona shows moderate increase in black population from 125000 to 250000, while
white populate shows significant increase from 1500000 to 2000000 points. California shows an increase from 1350000 to 1700000,
while white population drops from 12000000 to 10000000. This graph is important to consider for future changes in races
and the possibility to compare both race populations.

## Map


```{r echo=FALSE, warning=FALSE, message=FALSE}
map
```
I included this graph to compare black jail incarceration rate across the United states. 
The main purpose was to find what county has the
highest number of black people that are in the jail. The result shows that 
Los Angeles, California has the highest number of black people that are in the jail.
It shows that the approximate number is between 5000 and 7500
Another very high results show Miami Florida has around 2500 - 5000
black colored people in the jail
